---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

# Lab6

<!-- badges: start -->
<!-- badges: end -->

The goal of **Lab6** is to explore how algorithmic design and implementation affect computational performance in R.
We create an R package to study algorithms of different computational complexities and demonstrate ways to make R code faster.

In this lab, we will:

- bla

- bla

- bla...


## Installation

You can install the development version of **Lab6** from GitHub using:

```{r, eval = FALSE}
# don't already have devtools:
install.packages("devtools")

# install the package:
devtools::install_github("danielwalker/Lab6")
```


## Package vignette

This package includes a vignette titled **"lab report knapsack"**, which serves as the **main lab report** for **Lab 6**.  
The vignette contains the written answers, explanations, and performance comparisons between the three algorithms.

After installing and building the package, you can view the vignette using:

```{r, eval = FALSE}
browseVignettes("lab report knapsack")
```
