---
title: "lab report knapsack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab report knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(Lab6)
```

# Introduction

This vignette serves as the **lab report** for **Lab 6** in *Advanced R Programming*.

The purpose of this lab is to explore **algorithmic complexity** and **computational performance** using the **Knapsack problem**.


# The Knapsack Package

The **Knapsack problem** is a classic *discrete optimization problem*.  
We have a knapsack that can carry a limited weight `W`, and we want to fill it with `n` items —  
each item `i` having a weight `wᵢ` and a value `vᵢ`.  
The goal is to **maximize the total value** of the selected items without exceeding the weight limit.

Formally:

$$
\text{maximize} \quad \sum_{i=1}^n v_i x_i \\
\text{subject to} \quad \sum_{i=1}^n w_i x_i \le W, \quad x_i \in \{0,1\}
$$

This problem is **NP-hard**, meaning that it’s at least as hard as the hardest problems in NP.  
No polynomial-time algorithm is known for solving it optimally (this is an open Millennium Prize problem).  
For background, see the [Wikipedia page on the Knapsack problem](https://en.wikipedia.org/wiki/Knapsack_problem).

---

# Data Generation

The data for this lab is generated as follows.  
To create larger datasets, you can increase the value of `n`.  
Be careful with the random number generator version to ensure reproducibility.

```{r}
# Ensure consistent random number generation
RNGversion(min(as.character(getRversion()), "3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")

# Number of items
n <- 2000

# Generate weights and values
knapsack_objects <- data.frame(
  w = sample(1:4000, size = n, replace = TRUE),
  v = runif(n = n, 0, 10000)
)

# Display first few rows
head(knapsack_objects)
```

